---
output: github_document
format: gfm
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# arcgisplaces

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

arcgisplaces is an R package to interface with [ArcGIS Places Service](https://developers.arcgis.com/rest/places/).

> The places service is a ready-to-use location service that can search for businesses and geographic locations around the world. It allows you to find, locate, and discover detailed information about each place.

In order to use `{arcgisplaces}` you will need an ArcGIS Developers account. [Get started here.](https://developers.arcgis.com/documentation/mapping-apis-and-services/get-started/)

## Installation

You can install a binary of the development version of arcgisplaces from [r-universe](https://r-arcgis.r-universe.dev/arcgisplaces) with:

``` r
install.packages("arcgisplaces", repos = "https://r-arcgis.r-universe.dev")
```

You will also need the development version of [`{arcgisutils}`](https://github.com/R-ArcGIS/arcgisutils)

```r
if (!requireNamespace("pak")) install.packages("pak")
pak::pak("r-arcgis/arcgisutils")
```

### Building from source

Or, you can install the development version from [GitHub](https://github.com/r-arcgis/arcgisplaces). Note the development version requires an installation of Rust. See [rustup](https://rustup.rs/) for instructions to install Rust. 

```r
if (!requireNamespace("pak")) install.packages("pak")
pak::pak("r-arcgis/arcgisplaces")
```

## Usage

Finding places: 

- `near_point()`: search for places near a location.
- `within_extent()`: search for places within an extent.
- `place_details()`: get detailed information about the places returned from `near_point()` or `within_extent()`. 
  - Note: see `fields` for the possible attributes to return for place details.

Understanding categories: 

- `categories()`: find categories by name or ID.
- `category_details()`: get detailed information about the categories returned from `categories()`.

- Find place attributes such as name, address, description, opening hours, price ratings, user ratings, and social links.


## Examples

```{r include = FALSE}
library(sf)
library(pillar)

```

`{arcgisutils}` is needed for authentication. The Places API supports either using an API key via `auth_key()` or one generated via OAuth2 using either `auth_client()` or `auth_code()`. See the [Places API documentation](https://developers.arcgis.com/rest/places/#authentication) for more.

```{r }
library(arcgisutils)
library(arcgisplaces)

# Authenticate with a Developer Account API Key
token <- auth_key()
set_arc_token(token)
```

## Place search 

You can **search for places near a location** with `near_point()`.

```{r}
coffee <- near_point(x = -122.334, y = 47.655, search_text = "Coffee")
coffee
```

Locations are returned as an sf object with the place ID, the place name, distance from the search point, a character vector of categories. 


::: callout-tip

`arcgisplaces` will return an sf object, but the sf package is not required to work with the package. The `sf` print method will not be used unless the package is loaded. If package size is a consideration—i.e. deploying an app in a Docker container—consider using `wk` or `rsgeo`.

:::

Details for the places can be fetched using `place_details()`. The possible fields are [documented online](https://developers.arcgis.com/rest/places/place-id-get/#requestedfields) as well as contained in the exported vector `fields`. Because pricing is dependent upon which fields are requested, it is a required argument. 

To get the add `requested_fields = "hours"`. Note, that the other possible fields will still be present in the result, but completely empty. 

```{r}
details <- place_details(
  coffee$place_id,
  requested_fields = "rating",
  .progress = FALSE # remove progress bar
)

details[c("price", "user")]
```

Or, you can search for places within a bounding box using `within_extent()`. This could be quite handy for searching within current map bounds, for example. 

```{r}
bakeries <- within_extent(
  -70.356, 43.588, -70.176, 43.7182,
  category_id = "13002"
)

bakeries[c("name")]
```
